$\XX$ dubious ways of estimating $\rr$

### 2018-07-03 paper/draft_v7/draft_v7.tex
finally on git

Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease.  These methods are the delta method and posterior distribution.

@@

\textit{With the many different ways to estimate $\rr$, we want to compare them all to see which ones give more accurate and reliable estimates.}  As there is no way to know the ``true'' $\rr$ of an actual disease we compare our methods by fitting them to simulated data sets and look at the resulting $\rr$ estimates and standard errors. Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value. We vary the conditions under which the SIR data is generated to see how the results from the estimators change. We also look at data from different functional models to see how the estimators perform under different circumstances.

%% What do you mean by describing the standard standard deviations?
%% SG: delta_t is normally distributed with sd sigma and similarly for other methods.  How and why did we pick the numbers that we did?

\textit{To see the effects of different conditions for the data, we change one of the starting conditions while holding all of the other ones constant. We try changing two conditions: the starting number of susceptible and infected people in the population (holding the total population constant) and the variance of the errors added on to the The starting population sizes for the $X$ and $Y$ are set to ($X_0$, $Y_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant. The variance of the added errors by scaling them to: ($\sigma_X$, $\sigma_Y$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($X_0$, $Y_0$) = (99000, 1000). The errors start very small but become very large relative to the actual compartment sizes; this allows us to more clearly see the effects of the changes.}

@@

%% Talk about top 3-5 methods
%%%%%% Honestly, all the methods seem to have flaws in one way or another - maybe point out why?
%% Talk about methods that really miss
%% Talk about the standard error reports
%% Talk about 4 categories.  Does anyone do worse/better?
%% What about the one where we cannot estimate std
%% What happens when SD is large?

@@

\subsubsection{SIR Data}

\textit{The datasets from the SIR model are simulated from the ODEs: \begin{align}
	\systeme{\frac{dX}{dt} = -\frac{\beta XY}{N}, \frac{dY}{dt} = \frac{\beta XY}{N} - \gamma Y, \frac{dZ}{dt} = \gamma Z}.
	\end{align}
}

We use arbitrarily selected the parameters to generate the time series data set from the SIR model. We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated. The baseline SIR data are simulated with the following conditions:

@@

	$\beta$ = 0.006, $\gamma$ = 0.003; $\rr$ = 2

@@

	Starting X size ($X_0$): 99950

	Starting Y size ($Y_0$): 50

	Starting Z size: 0

@@

%% Is this enough detail on describing the initial S, I, and R, or are we looking for something more specific?

\textbf{Our baseline initial size for $Y_0$ is 0.05\% of the total population in our simulated environment, which is fairly large compared to the starting point for the data in most epidemics. However, I/we felt that this was necessary in order to see the differences between the models.}

We then generate random errors that are added to the X and Y compartments at each time step in order to simulate some form of randomness. \textbf{The sum of these was then subtracted from the Z compartment in order to keep the total population in the system constant.} We generated data sets with two different types of errors: Gaussian and autoregressive. \textbf{The Gaussian errors are independently distributed with mean 0 and constant variance for the entire length of the time series. The autoregressive errors are generated from a first-order process of the form: $$ \epsilon_{t} = \rho * \epsilon_{t-1} + \delta_{t}$$, where $\rho$ is fixed throughout and $\delta_{t} \sim N(0, \sigma)$ with constant $\sigma$. The values of the variances are chosen somewhat arbitrarily, although they are intended to be fairly small at first.}

We also add a monotonicity adjustment to the X and Z compartments with both of the error models. \textbf{This is done using the Pool Adjacent Violators Algorithm (PAVA); the algorithm checks if the observation at each time step is greater than or equal to the previous one, and changes the value of both observations to be equal to the mean of the two if the condition is not fulfilled.} This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM). For our baseline data, we have ($\sigma_X, \sigma_Y$) = (10, 1). We simulate 5 different baseline sets to see how consistent each of the models are.

\subsubsection{Other Models}
We simulated data from 6 other models besides the classic SIR model. 5 of these were polynomial linear regression models, from order 0 to 4 with respect to time/day. The coefficients used to generate data from the linear regression models were originally chosen by taking the approximate coefficient estimates from fitting the linear regression models to classic SIR data. This does not necessarily guarantee that the true $\rr$ will be 1.2, especially for the 0th and 1st order linear regression models. What we are interested in for these cases is to see if the models still give reasonable estimates for $\rr$. The other model is the linear SIR model, which is a three compartment model with the following ODEs:

@@

 The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2.

 \textbf{Random error is added to the data generated here as well, using the same 4 error models as above. The other specifications for these datasets are the same as for the baseline classic SIR data, namely the variance of the added errors and starting compartment sizes.}

\subsubsection{Variance Estimation}

\textbf{The method for calculating the standard errors for $\hat{\rr}$ depend on the model used. The following models take the standard error directly from the regression methods used to estimate $\rr$:}

	\begin{itemize}
		\item SIR LMA
		\item SIR LMAT
		\item SIR Log Linear
		\item SIR MC
		\item Exponential Growth
		\item Initial Growth
	\end{itemize}

\textbf{The ones that are calculated using the delta method:}

	\begin{itemize}
		\item-SIR LS
		\item SIR ReLS
		\item SEIR
		\item MSEIR
		%% \item AMP
		\item NGM
		\item Final Size
	\end{itemize}

\textbf{Standard errors for the maximum likelihood and branching process models are calculated using the negative reciprocal of the Fisher information. The sequential Bayes model uses the standard deviation from the posterior distribution. We use the jackknife method to estimate the standard error of the SIR smoothed max and the SIR smoothed IPR methods. The jackknife calculates the standard error using the standard deviation of estimates of $\rr$ after leaving out a single data point.}

\textbf{Some of the models use only one single point in time to estimate $\rr$ and therefore do not seem to have an innate source of variance. We currently do not have standard errors for the SIR Max and SIR IPR models.}

\subsection{Results}

\subsubsection{Survival Function}

%% How do we want to estimate this?????????
%% SG:
%% 3.1.1:  So I think this one we should mention there is no direct way to estimate this from the simple data we have.  Alternatively, you could do a parameter sweep with different beta distributions for F(a) and beta for b(a).
%% 3.1.2:  So assuming the population is homogeneous, k = N, D = gamma^{-1}.  So we have to estimmate $b$ from the data.  So I would average over (change in % susceptible ) / old # infected and use that as $b$.
%% 3.1.3:  No clue.

\subsubsection{Compartment Models}

\paragraph{SIR Models}
In general, the models derived from the SIR ordinary differential equations are relatively accurate. Most estimates are within the range of $\pm$ 0.2 of the true $\rr$ for the baseline SIR datasets. The magnitude of the standard errors on the baseline datasets for the different SIR models generally depend on the method used. The models that estimate $\rr$ using optimization or gradient descent, namely SIR LS and SIR ReLS, have standard errors on the order of $10^{-4}$ to $10^{-7}$. The standard errors are somewhat larger for the methods that use regression, on the order of $10^{-1}$ to $10^{-3}$.

The SIR models as a whole also do fairly well with changing the starting number of infected people. However, they tend not to do as well when the variance of the added errors is very large; in these situations, many of the models tend to have very large standard errors, sometimes on the order of $10^{1}$ or $10^{2}$.

The exceptions to the above are the linear, incidence-prevalence ratio, and Markov chain models. The estimates tend to be far from the true $\rr$, and they fluctuate greatly between the baseline data sets. The standard errors for these models (if they exist) also tend to be very large under some of the changing conditions.

Many of the SIR models do not work as well with the data that come from the polynomial linear models. Specifically, for the 0th and 1st order model datasets, the estimates can be unreasonably large or negative. The Log linear, Sequential Bayes, Smoothed max, and Max models do not have this problem.

\paragraph{Other Compartment Models}

All of these models use optimization to estimate the parameters that are used to calculate $\rr$. They tend to be fairly accurate and consistent on the baseline data, often as good as or better than those from the various SIR models. The SEIR models (least squares and next generation) do not handle the large added errors very well, as it tends to give ridiculously large results. The standard errors tend to be on the order of $10^{-2}$ to $10^{-3}$. However, as with the estimates above, the standard errors from the SEIR models can become unreasonably large when the variance of the added errors becomes large. Also, as with the SIR models, the $\rr$ estimates and standard errors from the 0th and 1st order polynomial linear model datasets are not very good, as the former is sometimes negative and the latter can be very large. The MSEIR model tends to do fairly well in all situations.

We have assumed for these models that the number of infections follows an exponential curve throughout the entire time frame of the data sets. This assumption certainly does not hold for the data from the 0th and 1st order linear models, and may not hold for some of the other data sets either.

These models are generally not as accurate as the compartment models above. Most of them, except for the initial size model, tend to underestimate $\rr$, with estimates between 0.8 and 1.0. The standard errors are all fairly small, on the order of magnitude from $10^{-1}$ to $10^{-3}$. This is larger than many of the compartment model standard errors. The main advantage that all of these models, except for the maximum likelihood model, seem to be more robust to changing conditions and functional forms of the data compared to the compartment models. The estimates and standard errors for $\rr$ appear to be reasonable in all cases, including for very large added errors and for the 0th and 1st order linear model data.

\subsubsection{Network Models}

The branching process model (the only one in this category so far) is not very accurate. In particular, for the SIR datasets, the $\rr$ estimates are less than 1, while the true $\rr$ is greater than 1. The standard errors, though, are relatively small, around the same order of magnitude as the SIR least squares models. Model estimates and standard errors are fairly consistent across the baseline datasets and changing starting compartment sizes. However, increasing the standard deviation of the added errors tends to have a somewhat noticeable impact on model results. The model also does not handle the 0th order polynomial linear data well.

Below, we compare the estimates and standard errors of $\rr$ given by each of the models for a few selected data sets. The forest plots below show the point estimates and 95\% confidence intervals of $\rr$ from each model for a particular data set. The true $\rr$ value is also shown by the vertical red line in the graphs.

All of the figures below show results from data sets that were simulated with added Gaussian errors. The plot on the top left shows the estimates from one of the baseline data sets. For the most part, the $\rr$ values range from approximately 0.8 to 1.4. We note that the range of values includes $\rr$ values of less than 1, meaning that some models predict that a disease is not spreading when it actually is. However, on the whole, the variation of estimates between models is smaller than we originally expected. The confidence intervals are very small relative to the estimates for the most part; exceptions to this include the SIR IPR and the Markov chain model as the exception.

The plot on the top right shows the estimates for the data set where $Y_0$ is increased to 100. The $\rr$ estimates generally seem to be further away from the true $\rr$ value compared to the results from the baseline data sets, and the overall . The confidence intervals are also more often than not larger than those in the baseline. In particular, the standard errors for some of the SIR models increase by several orders of magnitude.

The bottom left plot shows the estimates for the data set where the error added to the generated data is increased. For most of the models, the estimates of $\rr$ do not change very much compared to the baseline data set. The confidence intervals for this data set increase for some of the models, such as the SIR LMA, while they remain approximately the same as the baseline data sets for most of the others.

On the bottom right, we see the $\rr$ estimates for the case when disease prevalence grows linearly, as opposed to exponentially as seen in the data sets generated under SIR conditions, where the disease prevalence grows exponentially. Model performance here varies depending on how well the models the aforementioned linear growth. Some of the models, such as the SIR Markov chain and the IPR models, do not handle this case very well, giving much larger estimates and standard errors than are reasonable. On the other hand, models such as the SIR maximum and final size seem to be more flexible to different trajectories. [The error bars do not show up if they go outside of the x-axis limits - how do we fix this?]

\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{BaselineFP.jpeg}
\end{minipage}
\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{StartPopFP.jpeg}
\end{minipage}

\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{VarianceFP.jpeg}
\end{minipage}
\begin{minipage}{0.5\linewidth}
	\includegraphics[width=3in]{OtherFP.jpeg}
\end{minipage}

\section{Appendix}

Included here are the technical details for fitting the models in R, as well as tables of the results.

@@

For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $Y_0$. This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_Y$ at 1 no matter if $Y_0$ is 1 or 1000.

For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_X$ and $\sigma_Y$. However, it is interesting to note that the standard errors are increasing for the monotonic Gaussian error data, but decreasing for the autoregressive and monotonic autoregressive error data and relatively stable for the Gaussian non-monotonic data. It is not clear why this would be the case.

\begin{table}[H]
	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 1.965322 & 0.000869 & 1.983308 & 0.000882 & 2.005867 & 0.000893 & 1.994736 & 0.000884\\
			\hline
			Baseline 2 & 2.039608 & 0.000913 & 1.999245 & 0.000895 & 2.001217 & 0.000889 & 1.995960 & 0.000887\\
			\hline
			Baseline 3 & 2.005609 & 0.000901 & 1.966631 & 0.000873 & 2.016216 & 0.000901 & 1.997175 & 0.000887\\
			\hline
			Baseline 4 & 1.958037 & 0.000859 & 1.941109 & 0.000852 & 1.995544 & 0.000886 & 2.004107 & 0.000892\\
			\hline
			Baseline 5 & 2.050904 & 0.000926 & 1.964328 & 0.000866 & 2.009912 & 0.000896 & 1.990153 & 0.000881\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 1.689292 & 0.020794 & -4.167216 & 0.168853 & 1.976254 & 0.045081 & 1.688593 & 0.029680\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 2.038297 & 0.004597 & 2.207003 & 0.005165 & 2.037726 & 0.004596 & 1.883084 & 0.004049\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 2.009581 & 0.000449 & 2.006557 & 0.000448 & 1.999236 & 0.000445 & 1.995467 & 0.000444\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 2.002521 & 0.000046 & 2.001829 & 0.000046 & 1.999519 & 0.000046 & 2.000098 & 0.000046\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Different Variances, $X_0 = 99000$, $Y_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$\sigma_X = 10, \sigma_Y = 1$ & 1.927444 & 0.000059 & 1.927090 & 5.9e-05 & 1.925768 & 5.9e-05 & 1.926576 & 5.9e-05\\
			\hline
			$\sigma_X = 100, \sigma_Y = 10$ & 1.926786 & 0.000058 & 1.916571 & 5.9e-05 & 1.919778 & 5.9e-05 & 1.929422 & 5.9e-05\\
			\hline
			$\sigma_X = 300, \sigma_Y = 30$ & 1.994211 & 0.000062 & 1.976893 & 5.8e-05 & 1.930660 & 5.8e-05 & 1.920869 & 5.9e-05\\
			\hline
			$\sigma_X = 1000, \sigma_Y = 100$ & 1.749057 & 0.000050 & 1.736588 & 5.9e-05 & 2.003654 & 6.2e-05 & 1.897049 & 5.9e-05\\
			\hline
			$\sigma_X = 3000, \sigma_Y = 300$ & 1.992674 & 0.000063 & 1.681585 & 4.9e-05 & 1.959815 & 6.7e-05 & 1.770816 & 5.3e-05\\
			\hline
			$\sigma_X = 10000, \sigma_Y = 1000$ & -42.264180 & 0.002694 & 1.773388 & 3.1e-05 & 1.889188 & 6.6e-05 & 2.649010 & 6.3e-05\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Other Models, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			0th & -2.032207 & 0.244058 & 1.229963 & 0.069020 & 1.021369 & 0.056558 & 0.795460 & 0.331468\\
			\hline
			1st & 1.038117 & 0.000029 & 1.037889 & 0.000029 & 1.037902 & 0.000029 & 1.037887 & 0.000029\\
			\hline
			2nd & 1.198856 & 0.000045 & 1.198589 & 0.000045 & 1.198772 & 0.000045 & 1.198501 & 0.000045\\
			\hline
			3rd & 1.202405 & 0.000043 & 1.202098 & 0.000043 & 1.202350 & 0.000043 & 1.202240 & 0.000043\\
			\hline
			4th & 1.210597 & 0.000047 & 1.210597 & 0.000047 & 1.210194 & 0.000047 & 1.210304 & 0.000047\\
			\hline
			Linear SIR & 1.274346 & 0.000034 & 1.273739 & 0.000034 & 1.274229 & 0.000034 & 1.273950 & 0.000034\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_X$, $\sigma_Y$) increase, while they become less accurate for the monotonic values. The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $X_0$ and $Y_0$ are held constant.

@@

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Baseline Data, $X_0 = 99950, Y_0 = 50$, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			Baseline 1 & 3.226272 & 0.000000 & 1.984447 & 0.000294 & 2.006310 & 0.000298 & 1.993558 & 0.000294\\
			\hline
			Baseline 2 & 3.203012 & 0.000000 & 1.999589 & 0.000299 & 2.000868 & 0.000297 & 1.995630 & 0.000296\\
			\hline
			Baseline 3 & 4.180804 & 0.000733 & 4.180810 & 0.000724 & 5.043355 & 0.000908 & 1.997399 & 0.000296\\
			\hline
			Baseline 4 & 4.180819 & 0.000709 & 4.180816 & 0.000715 & 1.995237 & 0.000295 & 2.003821 & 0.000298\\
			\hline
			Baseline 5 & 4.180799 & 0.000741 & 4.180815 & 0.000717 & 2.008713 & 0.000300 & 4.180812 & 0.000720\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Different Starting Populations, 
		$\sigma_X = 10, \sigma_Y = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
			\hline
			$X_0 = 99999, Y_0 = 1$ & 5.899238 & 0.025009 & 5.898924 & 0.051433 & 3.394166 & 0.000021 & 3.217232 & 0.000016\\
			\hline
			$X_0 = 99990, Y_0 = 10$ & 3.162364 & 0.000001 & 3.170152 & 0.000002 & 3.131576 & 0.000001 & 3.108194 & 0.000001\\
			\hline
			$X_0 = 99900, Y_0 = 100$ & 4.564960 & 0.000404 & 2.005939 & 0.000150 & 1.998880 & 0.000149 & 1.994592 & 0.000148\\
			\hline
			$X_0 = 99000, Y_0 = 1000$ & 47.241714 & 0.002493 & 47.241714 & 0.002511 & 47.241714 & 0.002502 & 47.241714 & 0.002502\\
			\hline
		\end{tabular}
	\end{footnotesize}
\end{table}

>>>>

Additionally, we describe  interval estimates for the above methods for estimating $\rr$ in order to better answer the question of whether $\rr > 1$, which implies an outbreak of the disease.  These methods are the delta method, and posterior distribution.

@@

To compare the different models, we fit each of them to simulated data sets and look at the resulting $\rr$ estimates and standard errors. Our main comparisons are done with data generated from the SIR model with random error added so that we can compare model estimates to a known true $\rr$ value. We vary the conditions under which the SIR data is generated to see how the results from the estimators change. We also look at data from different functional models to see how the estimators perform under different circumstances.

@@

%% Talk about top 3-5 methods
%% Talk about methods that really miss
%% Talk about the standard error reports
%% Talk about 4 categories.  Does anyone do worse/better?
%% What about the one where we cannot estimate std
%% What happens when SD is large?

@@

\subsubsection{SIR Data}
We use arbitrarily selected parameters to generate the time series data set from the SIR model. We first generate "baseline" data from which we can make comparisons by changing the  conditions under which the data were generated. The baseline SIR data are simulated with the following conditions:

@@

	$\beta$ = 0.06, $\gamma$ = 0.05; $\rr$ = 1.2

@@

	Starting S size ($S_0$): 99950

	Starting I size ($I_0$): 50

	Starting R size: 0

@@

We also generate random errors that are added to the S and I compartments at each time step. We had two types of errors: independent Gaussian and autoregressive. We also added a monotonicity adjustment to the S and R compartments with both of the error models. This gives us a total of four error models: autoregressive (Auto), autoregressive monotonic (AutoM), Gaussian (Norm), and Gaussian monotonic (NormM). For our baseline data, we have ($\sigma_S, \sigma_I$) = (10, 1). We simulate 5 different baseline sets to see how consistent each of the models are.

To see the effects of different conditions for the data, we change the starting population sizes for the $S$ and $I$ are to ($S_0$, $I_0$) =  (99999, 1), (99990, 10), (99900, 100), and (99000, 1000), holding all other conditions constant. We also change the variance of the added errors by scaling them to: ($\sigma_S$, $\sigma_I$) = (10, 1), (100, 10), (500, 50), (1000, 100), (2000, 200), (5000, 500), (10000, 1000), (15000, 1500); these are done with ($S_0$, $I_0$) = (99000, 1000) and all other conditions the same as the baseline.

\subsubsection{Other Models}
We simulated data from 6 other models besides the classic SIR model. 5 of these were linear regression models, from order 0 to 4 with respect to time/day. The coefficients used to generate data from the linear regression models were originally chosen by taking the approximate coefficient estimates from fitting the linear regression models to classic SIR data. This does not necessarily guarantee that the true $\rr$ will be 1.2, especially for the 0th and 1st order linear regression models. The other model is the linear SIR model, which is a three compartment model with the following ODEs:

@@

 The data from the linear SIR model were simulated with parameters $\beta$ = 0.06 and $\gamma$ = 0.05, such that $\rr$ = 1.2. The other specifications for the data simulated from the other models (the variance of the errors and starting compartment sizes) are the same as for the baseline classic SIR data.

@@

For the changing compartment sizes, we see that the standard errors are much larger for smaller values of $I_0$. This might be because the variance of the I compartment is much larger relative to the size of the compartment, since we hold $\sigma_I$ at 1 no matter if $I_0$ is 1 or 1000.

For the changing error variances, we see that the estimates for $\rr$ become less accurate for increasing $\sigma_S$ and $\sigma_I$. However, it is interesting to note that the standard errors are increasing for the monotonic Gaussian error data, but decreasing for the autoregressive and monotonic autoregressive error data and relatively stable for the Gaussian non-monotonic data. It is not clear why this would be the case.

\begin{table}[H]
	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Baseline Data, $S_0 = 99950, I_0 = 50$, 
		$\sigma_S = 10, \sigma_I = 1$}
	\begin{footnotesize}
		\hskip -1cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		Baseline1 & 1.199737 & 0.000476 & 1.199557 & 0.000476 & 1.199737 & 0.000475 & 1.199737 & 0.000475\\
		\hline
		Baseline2 & 1.199737 & 0.000475 & 1.200380 & 0.000481 & 1.200376 & 0.000482 & 1.200376 & 0.000482\\
		\hline
		Baseline3 & 1.200615 & 0.000485 & 1.199557 & 0.000476 & 1.200380 & 0.000481 & 1.200298 & 0.000479\\
		\hline
		Baseline4 & 1.200376 & 0.000481 & 1.200061 & 0.000479 & 1.200380 & 0.000481 & 1.200602 & 0.000484\\
		\hline
		Baseline5 & 1.200376 & 0.000481 & 1.200376 & 0.000482 & 1.200298 & 0.000479 & 1.200602 & 0.000484\\
		\hline
	\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Different Starting Populations, 
		$\sigma_S = 10, \sigma_I = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		$S_0 = 99999, I_0 = 1$ & 1.214968 & 0.095795 & 1.236988 & 0.096579 & 1.218704 & 0.099065 & 1.218704 & 0.096508\\
		\hline
		$S_0 = 99990, I_0 = 10$ & 1.219334 & 0.006658 & 1.219334 & 0.006854 & 1.219596 & 0.017843 & 1.219596 & 0.017847\\
		\hline
		$S_0 = 99900, I_0 = 100$ & 1.199311 & 0.000166 & 1.200216 & 0.000170 & 1.200336 & 0.000170 & 1.200336 & 0.000170\\
		\hline
		$S_0 = 99000, I_0 = 1000$ & 1.199807 & 0.000029 & 1.199807 & 0.000029 & 1.200160 & 0.000029 & 1.200160 & 0.000029\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model, 
		Different Variances, $S_0 = 99000$, $I_0 = 1000$}
	\begin{footnotesize}
		\hskip -1.7cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Error & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		$\sigma_S = 10, \sigma_I = 1$ & 1.200061 & 2.9e-05 & 1.200005 & 2.9e-05 & 1.199951 & 2.9e-05 & 1.199916 & 2.9e-05\\
		\hline
		$\sigma_S = 100, \sigma_I = 10$ & 1.200317 & 2.9e-05 & 1.200427 & 2.9e-05 & 1.199963 & 2.9e-05 & 1.199734 & 2.9e-05\\
		\hline
		$\sigma_S = 500, \sigma_I = 50$ & 1.201642 & 2.8e-05 & 1.201871 & 2.8e-05 & 1.199996 & 2.9e-05 & 1.198872 & 2.9e-05\\
		\hline
		$\sigma_S = 1000, \sigma_I = 100$ & 1.203259 & 2.8e-05 & 1.203811 & 2.8e-05 & 1.199985 & 2.9e-05 & 1.197928 & 2.9e-05\\
		\hline
		$\sigma_S = 2000, \sigma_I = 200$ & 1.206447 & 2.7e-05 & 1.207566 & 2.7e-05 & 1.199849 & 2.9e-05 & 1.195612 & 3.0e-05\\
		\hline
		$\sigma_S = 5000, \sigma_I = 500$ & 1.215906 & 2.5e-05 & 1.218773 & 2.5e-05 & 1.199656 & 2.9e-05 & 1.189382 & 3.2e-05\\
		\hline
		$\sigma_S = 7500, \sigma_I = 750$ & 1.223554 & 2.3e-05 & 1.228037 & 2.4e-05 & 1.199579 & 2.9e-05 & 1.183914 & 3.4e-05\\
		\hline
		$\sigma_S = 10000, \sigma_I = 1000$ & 1.230992 & 2.2e-05 & 1.237142 & 2.2e-05 & 1.199381 & 2.9e-05 & 1.178701 & 3.6e-05\\
		\hline
		$\sigma_S = 15000, \sigma_I = 1500$ & 1.245535 & 1.9e-05 & 1.255188 & 2.0e-05 & 1.199137 & 2.9e-05 & 1.168520 & 4.1e-05\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIR Model,
		Other Models, $S_0 = 99950, I_0 = 50$, 
		$\sigma_S = 10, \sigma_I = 1$}
	\begin{footnotesize}
		\hskip -1cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		0th & 0.005468 & 0.000745 & -0.034869 & 0.000764 & -0.034869 & 7.4e-04 & 0.034869 & 0.000673\\
		\hline
		1st & 1.038004 & 0.000029 & 1.037900 & 0.000029 & 1.037826 & 2.9e-05 & 1.037826 & 0.000029\\
		\hline
		2nd & 1.199663 & 0.000045 & 1.198627 & 0.000045 & 1.199663 & 4.5e-05 & 1.198949 & 0.000045\\
		\hline
		3rd & 1.202909 & 0.000043 & 1.202282 & 0.000043 & 1.202414 & 4.3e-05 & 1.202414 & 0.000043\\
		\hline
		4th & 1.211092 & 0.000047 & 1.211092 & 0.000047 & 1.209479 & 4.7e-05 & 1.211092 & 0.000047\\
		\hline
		Linear SIR & 1.274529 & 0.000034 & 1.273768 & 0.000034 & 1.275057 & 3.4e-05 & 1.274745 & 0.000034\\
		\hline
	\end{tabular}
\end{footnotesize}
\end{table}

@@

The $\rr$ estimates for the data with non-monotonic errors become more accurate as ($\sigma_S$, $\sigma_I$) increase, while they become less accurate for the monotonic values. The standard errors are also positively correlated with the estimate of $\rr$ as long as all other parts, i.e., $S_0$ and $I_0$ are held constant.

@@

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Baseline Data, $S_0 = 99950, I_0 = 50$, 
		$\sigma_S = 10, \sigma_I = 1$}
	\begin{footnotesize}
		\hskip -1cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		Baseline 1 & 1.200196 & 1e-06 & 1.199942 & 1e-06 & 1.200108 & 1e-06 & 1.199965 & 1e-06\\
		\hline
		Baseline 2 & 1.199574 & 1e-06 & 1.199733 & 1e-06 & 1.199922 & 1e-06 & 1.200219 & 1e-06\\
		\hline
		Baseline 3 & 1.199909 & 1e-06 & 1.200416 & 1e-06 & 1.199851 & 1e-06 & 1.200009 & 1e-06\\
		\hline
		Baseline 4 & 1.199735 & 1e-06 & 1.199829 & 1e-06 & 1.199925 & 1e-06 & 1.199856 & 1e-06\\
		\hline
		Baseline 5 & 1.199849 & 1e-06 & 1.199683 & 1e-06 & 1.199943 & 1e-06 & 1.199900 & 1e-06\\
		\hline
	\end{tabular}
	\end{footnotesize}
\end{table}

@@

	\caption{$\rr$ Estimates and Std. Errs, SIRREPAR Model,
		Different Starting Populations, 
		$\sigma_S = 10, \sigma_I = 1$}
	\begin{footnotesize}
		\hskip -1.7cm
	\begin{tabular}{l|r|r|r|r|r|r|r|r}
		\hline
		Data Set & Auto Est & Auto SE & AutoM Est & AutoM SE & Norm Est & Norm SE & NormM Est & NormM SE\\
		\hline
		$S_0 = 99999, I_0 = 1$ & 1.209265 & 3.7e-05 & 1.198242 & 3.5e-05 & 1.193162 & 3.4e-05 & 1.200995 & 3.5e-05\\
		\hline
		$S_0 = 99990, I_0 = 10$ & 1.202026 & 4.0e-06 & 1.198515 & 4.0e-06 & 1.199652 & 4.0e-06 & 1.199716 & 4.0e-06\\
		\hline
		$S_0 = 99900, I_0 = 100$ & 1.199895 & 1.0e-06 & 1.199895 & 1.0e-06 & 1.199976 & 1.0e-06 & 1.200103 & 1.0e-06\\
		\hline
		$S_0 = 99000, I_0 = 1000$ & 1.200089 & 0.0e+00 & 1.200008 & 0.0e+00 & 1.199992 & 0.0e+00 & 1.199985 & 0.0e+00\\
		\hline
	\end{tabular}
	\end{footnotesize}
\end{table}
